### 2022-11-25 draft.tex
Deploying to gh-pages from @ graciellehigino/ms_range_interactions@44cf0d87cc5a83ee9702e8b4b9bbcfb9c91ec7a3 🚀

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to detect areas of potential misestimation of species'
distribution data (more precisely range maps) based on interaction data,
considering the basic assumption that predators can only be present in
regions where they are connected to at least one herbivore - and thus
indirectly connected to primary producers. We used a Serengeti food web
dataset (Baskerville et al. 2011) (which comprises carnivores,
herbivores, and plants from Tanzania) to demonstrate how a mismatch
between occurrence and interaction data can highlight significant
uncertainty areas in IUCN range maps. Finally, we add the GBIF
occurrence points for the Serengeti species to the investigation,
discuss the mechanisms that can lead to the lack of agreement between
data, and build from that a vision for the next steps, reinforcing the
importance of geographically explicit interaction data.

@@

Organisms cannot persist unless they are directly or indirectly
connected to a primary producer within their associated food web (Power
1992). Therefore, the range of a predator (omnivore or carnivore)
depends on the overlapping ranges of its preys. If sections of a
predator's range does not overlap with at least one of its prey it will
become disconnected from primary producers, and therefore we would not
expect the predator to occur in this area. This mismatch can be the
result of different mechanisms, like the misestimation of both the
predator's and the preys' ranges (Ladle and Hortal 2013; Rondinini et
al. 2006), taxonomic errors (Isaac, Mallet, and Mace 2004; Ladle and
Hortal 2013), or the lack of information about trophic links (i.e., the
lack of connection between the ranges of a predator and a primary
producer may be due a third species we don't know is connected to both).
Thus, given that herbivores are the main connection between plant
resources (directly limited by environmental conditions) and predators
(Dobson 2009; Scott et al. 2018), here we adjusted the ranges of
predators based on a simple rule: we removed any part of a predator's
range that did not intersect with the range of at least one prey
herbivore species. So, unless the range of the predator overlapped with
at least one prey item, which in turn is directly connected to a primary
producer (plants), we removed that section of the predator's range.
Finally, we calculated the difference in range size between the original
IUCN ranges and those adjusted based on species interaction data.

\hypertarget{data}{%
\subsection{Data}\label{data}}

@@

We compiled IUCN range maps for the 32 species included in the metaweb
from the Spatial Data Download portal
(www.iucnredlist.org/resources/spatial-data-download), which we
rasterized at 10 arc-minute resolution (\textasciitilde18 km at the
equator). We restricted the rasters a spatial extent comprised between
latitudes 35°S and 40°N and longitudes 20°W and 55°E. We then combined
interaction data from the metaweb and cooccurrence data generated from
species ranges to create networks for each raster pixel. This generated
a total of 84,244 pixel-level networks. These networks describe
potential predation, not actual interactions: the former is derived
information from the metaweb, and the latter is contingent on the
presence of herbivores.

>>>>

It has been demonstrated that the agreement between range maps and point
data varies geographically (Hurlbert and Jetz 2007; Hurlbert and White
2005; Ficetola et al. 2014). Adding ecological interaction data to this
comparison might help to elucidate where these (dis)agreements are more
likely to be true and which dataset better represent the actual
distribution of species. In this context, we elaborate a method that
allows us to detect areas of potential misestimation of species'
distribution data (more precisely range maps) based on interaction data.
This method is based on the assumption that organisms cannot persist in
an area unless they are directly or indirectly connected to a primary
producer within their associated food web (Power 1992). Thus, given that
herbivores are the main connection between plant resources (directly
limited by environmental conditions) and predators (Dobson 2009; Scott
et al. 2018), the range of a predator (omnivore or carnivore) depends on
the overlapping ranges of its herbivore preys. If sections of a
predator's range does not overlap with at least one of its prey it will
become disconnected from primary producers, and therefore we would not
expect the predator to occur in this area.

This mismatch can be the result of different mechanisms, like the
misestimation of both the predator's and the preys' ranges (Ladle and
Hortal 2013; Rondinini et al. 2006), taxonomic errors (Isaac, Mallet,
and Mace 2004; Ladle and Hortal 2013), or the lack of information about
trophic links (i.e., the lack of connection between the ranges of a
predator and a primary producer may be due a third species we don't know
is connected to both). Here we discuss the mechanisms that can lead to
the lack of agreement between data, and build from that a vision for the
next steps, reinforcing the importance of geographically explicit
interaction data.

@@

We indentified areas of uncertainty on the ranges of predators based on
a simple rule: we removed any part of a predator's range that did not
intersect with the range of at least one prey herbivore species, which
in turn is directly connected to a primary producer (plants). To do
that, we used a Serengeti food web dataset (Baskerville et al. 2011)
(which comprises carnivores, herbivores, and plants from Tanzania) and
its species ranges from IUCN. Then, we calculated the difference in
range sizes between the original IUCN ranges of predators and those
without the areas where they would be alone, based on species
interaction data. Finally, we added the GBIF occurrence points for the
Serengeti species to investigate whether the results would be different
if we used another source of distribution data. \#\# Data

@@

We compiled IUCN range maps for the 32 species included in the metaweb
from the Spatial Data Download portal
(www.iucnredlist.org/resources/spatial-data-download), which we
rasterized at a 0.5 degrees resolution (\textasciitilde50 km at the
equator). We restricted the rasters to a spatial extent comprised
between latitudes 35°S and 40°N and longitudes 20°W and 55°E. We then
combined interaction data from the metaweb and cooccurrence data
generated from species ranges to create networks for each raster pixel.
This generated a total of 11,308 pixel-level networks. These networks
describe potential predation, not actual interactions: the former is
derived information from the metaweb, and the latter is contingent on
the presence of herbivores.
